---
title: "In-class Exercise 3"
author: "Fan Syong Yue"
date: 08/26/24
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
---

# Packages

As `maptools` was removed from the CRAN repository, we can only install it through an archived link.

```{r}
#| eval: false
install.packages("maptools", repos = "https://packagemanager.posit.co/cran/2023-10-13")
```


We can now load the modules required

```{r}
pacman::p_load(tmap, sf, raster, spatstat, maptools, sp, tidyverse)
```

# Data Wrangling
we will be using the following datasets:

-   [Master Plan 2014 Subzone Boundary Data](https://beta.data.gov.sg/datasets/d_d14da225fccf921049ab64238ff473d9/view)
-   [Childcare Services](https://data.gov.sg/collections/2015/view) (GeoJSON)
-   CostalOutline from SLA

We first use `st_read()` to import all our into variables `mpsz_sf`,`childcare_sf`, `sg_sf`

```{r}
mpsz_sf <- st_read(dsn = "data", layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
childcare_sf <- st_read("data/child-care-services-geojson.geojson") %>%
  st_transform( crs = 3414)
```

```{r}
sg_sf <- st_read(dsn = "data", layer = "CostalOutline")
```

```{r}
acled_sf <- read_csv("data/Myanmar/ACLED_Myanmar.csv") %>%
  st_as_sf(coords = c(
    "longitude", "latitude"),
    crs=4326) %>%
  st_transform(crs = 32647) %>%
  mutate(event_date = dmy(event_date))
```
## Working with `st_union()`
We want to derive the coastal outline sf tibble data.frame

```{r}
sg_sf <- mpsz_sf %>%
  st_union()
```

