---
title: "Hands-on Exercise 2: Thematic Mapping and GeoVisualisation"
author: "Fan Syong Yue"
date: 08/24/24
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
---

# Overview

------------------------------------------------------------------------

Thematic mapping utilises map symbols to highlight geographic features that are not naturally visible. Examples include population, temperature and crime rate.

Geovisualisation, however, uses graphics to make a place, phenomenon, or process visible, tapping into our powerful spatial cognition and eye-brain vision system.

We will learn more about how to plot functional and truthful choropleth maps using the `tmap` package.

# Setup

## Packages

We will be using the following packages aside from `tmap`

-   tidyverse
    -   readr - importing delimited text files
    -   tidyr - tidying data
    -   dplyr - wrangling data
-   sf - handling geospatial data

```{r}
pacman::p_load(sf,tmap,tidyverse)
```

## Data

The following datasets will be used

-   [Master Plan 2014 Subzone Boundary Data](https://beta.data.gov.sg/datasets/d_d14da225fccf921049ab64238ff473d9/view)
-   [Singapore Residents by Planning Are / Subzone, Age Group, Sex, and Dwelling](https://www.singstat.gov.sg/-/media/files/find_data/population/statistical_tables/respopagesextod2011to2020.ashx)

We first use `st_read()` to import the geospatial data into a variable called `mpsz`

```{r}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
mpsz
```

We now import the aspatial data in an R dataframe `popdata`

```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

## Data Preparation

We first want to prepare a data table with year 2020 values with the variables PA, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY.

-   YOUNG: Age groups 0-4 until 20-24

-   ECONOMY ACTIVE: Age groups 25-29 until 60-64

-   AGED: Age group 65+

-   TOTAL: All age groups

-   DEPENDENCY: ratio between the YOUNG and AGED ages groups against the ECONOMY ACTIVE age group

### Data Wrangling

We will now use `pivot_wider()` from `tidyr` along with `mutate()`, `filter()`, `group_by()` and `select()` from `dplyr`

```{r}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from = AG, values_from = POP) %>%
  mutate(`YOUNG` = rowSums(.[3:6]) + rowSums(.[14])) %>%
  mutate(`ECONOMY ACTIVE` = rowSums(.[7:13])+rowSums(.[15])) %>%
  mutate(`AGED` = rowSums(.[16:21])) %>%
  mutate(`TOTAL` = rowSums(.[3:21])) %>%
  mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, `ECONOMY ACTIVE`, `AGED`, `TOTAL`, `DEPENDENCY`)
```

### Joining attribute and geospatial data

As the values in the PA and SZ fields are made up of upper- and lowercase characters, we will need to convert them all into uppercase to be used when joining with `SUBZONE_N` and `PLN_AREA_N`.

```{r}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = list(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

We can now properly join the data using the standardised names using `left_join()`

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020, by = c("SUBZONE_N" = "SZ"))
```

We want to now save this into a file

```{r}
write_rds(mpsz_pop2020, "data/mpszpop2020.rds")
```

# Choropleth Mapping Geospatial Data Using `tmap`

Choropleth mapping involves the symbolisation of enumeration units, such as countries, provinces, states, counties or census units, using area patterns or graduated colors.

Two approaches can be used to prepare thematic map using `tmap`, they are: - Plotting quickly using `qtm()` - Plotting highly customisable thematic maps by using `tmap` elements

## Plotting a choropleth map using `qtm()`

Using `qtm()` from the `tmap` package us the easiest and quickest way to draw a choropleth map. It provides a good and concise default vistualisation that covers a large number of cases.

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, fill = "DEPENDENCY")
```

From the code above, we set `tmap_mode("plot")` which provides us with a static map. If we required an interactive mode, "view" in replace of "plot" within `tmap_mode()` would be required. We should also note of the `fill` argument which selects which attribute we are mapping.

While useful for quickly producing a choropleth map, `qtm()` is not without disadvantages. Aesthetics of individual layers are hard to control.

## Creating a choropleth map by using `tmap`’s elements

If we wanted to draw a high quality cartographic choropleth map, we want to use `tmap`'s drawing elements.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          title = "Dependency Ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by Planning Subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45,
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type = "8star", size = 2) +
  tm_scale_bar() + tm_grid(alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS",
             position = c("left", "bottom"))
```

Now, let us break down the individual subfunctions used to make the above map

### Base map

We use `tm_shape()` to define our input data and `tm_polygons()` to draw out the map.

```{r}
tm_shape(mpsz_pop2020) + tm_polygons()
```

### Draw a Choropleth Map with `tm_polygons()`

We now want to specify the variable we want to use for drawing the choropleth map

```{r}
tm_shape(mpsz_pop2020) + tm_polygons("DEPENDENCY")
```

We should note that the default binning used to draw the choropleth map is called "pretty", the default colour scheme used is `YlOrRd` from ColorBrewer and by default, missing values are shaded in grey.

Furthermore, `tm_shape()` is in fact a wrapper function of `tm_fill()` and `tm_border()`. `tm_fill()` shades the polygons using the default colour scheme while `tm_border()` adds the borders of the shapefile onto the choropleth map.

### Drawing choropleth maps using `tm_fill()` and `tm_border()`

Let's first see how tm_fill works when used alone

```{r}
tm_shape(mpsz_pop2020) + tm_fill("DEPENDENCY")
```

Notice how the planning subzones are shared according to the respective dependency values? We can solve this using `tm_borders()`

```{r}
tm_shape(mpsz_pop2020) + tm_fill("DEPENDENCY") + tm_borders(lwd = 0.1, alpha = 1)
```

The *alpha* argument is used to define transparency using a value between 0 (transparent) and 1 (opaque). By default, the alpha value of the col is used (normally 1).

Besides *alpha*, there are three other arguments for tm_borders():

-   *col* -\> border color,

-   *lwd* -\> borderline width, default 1,

-   *lty* -\> borderline type, default "solid' .

## Data Classification using `tmap`

`tmap` provides us with ten classification methods: *fixed*, *sd*, *equal*, *pretty* (default), *quantile*, *kmeans*, *hclust*, *bclust*, *fisher*, and *jenks*.
These can be specified in `tm_fill()` or `tm_polygons()`

### Kmeans example with 5 classes

```{r}
tm_shape(mpsz_pop2020) + tm_fill("DEPENDENCY", n = 5, style = "kmeans") + tm_borders(alpha = 0.5)
```

### sd example with 5 classes
```{r}
tm_shape(mpsz_pop2020) + tm_fill("DEPENDENCY", n = 5, style = "sd") + tm_borders(alpha = 0.5)
```

## Choropleth Maps and Custom breaks
With built-in styles, category breaks are computed internally. To override defaults, we can set breakpoints using the `breaks` argument in `tm_fill()`. We should also not that `tmap` breaks have a maximum and minimum. Thus, to retrieve n categories, n+1 elements must be specified in the `breaks` argument in increasing order.

Before we explore more on this, let's take a look at some statistics of the `DEPENDENCY` field.

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

Using the above results as reference, we set the breakpoints at .6, .7, .8, and .9. With the addition of the min and max breakpoints, the *breaks* vector is c(0, 0.6, 0.7, 0.8, 0.9, 1).

```{r}
tm_shape(mpsz_pop2020) + tm_fill("DEPENDENCY", breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00, 19)) + tm_borders(alpha = 0.5)
```

## Colour Scheme

`tmap` supports custom colour ramps defined by the user, or from prebuilt sets from the `RColorBrewer` package

We can try this out using the `Blues` colour palette by setting the `palette` argument in tm_fill()

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```

We can also reverse the colour shading by adding a `-` prefix

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```

## Map Layout

Map layout refers to the combination of all map elements into a cohensive map. Map elements include among others the objects to be mapped, the title, the scale bar, the compass, margins and aspects ratios.

### Map Legend
Within `tmap` there are several options to modify a legend
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "jenks",
          palette = "Blues",
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby Planning Subzone",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45,
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

### Map Style
`tmap` allows a wide variety of layout settings to be changed. They can be called by using `tmap_style()`.

Let's first try out the "classic" style

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Blues") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

### Cartographic Furniture
Compass, scale bar, and grid lines can also be added using `tmap`.
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby Planning Subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.outside.size = 0.2,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star",
             size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1,
          alpha = 0.2) +
  tm_credits("sajdad",
             position = c("left", "bottom"))
```

To reset styling, we can use `tmap_style()`
```{r}
tmap_style("white")
```

## Small Multiple Maps A.K.A Facet Maps
These maps are composed of many maps arrange side-by-side, and sometimes stacked vertically. Small multiple maps enable the visualisation of how spatial relationships change with respect to another variable, such as time.

There are 3 ways using `tmap` that we can plot these maps:
-   assigning multiple values to at least one of the aesthetic arguments
-   defining a group-by variable in `tm_facets()`
-   creating multiple stand-alone maps with `tmap_arrange()`

### Using multiple values
In this example, small multiple choropleth maps are created by defining `ncols` in `tm_fill()`
```{r}
tm_shape(mpsz_pop2020) +
  tm_fill(c("YOUNG", "AGED"),
          style = "equal",
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5)
```

In this example, small multiple choropleth maps are created by assigning multiple values to at least one of the aesthetic arguments
```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons(c("DEPENDENCY", "AGED"),
              style = c("equal", "quantile"),
              palette = list("Blues", "Greens")) +
  tm_layout(legend.position =  c("right", "bottom"))
```

### Group-by variable in in `tm_facets()`
In this example, multiple small choropleth maps are created by using `tm_facets()`.

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette =  "Blues",
          thres.poly = 0) +
  tm_facets(by="REGION_N",
            free.coords = TRUE,
            drop.units = TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"),
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

### Creating multiple stand-alone maps with `tmap_arrange()`
In this example, multiple small choropleth maps are created by creating multiple stand-alone maps with `tmap_arrange()`.

```{r}
youngmap <- tm_shape(mpsz_pop2020) + tm_polygons ("YOUNG", style = "quantile", palette = "Blues")
agedmap <- tm_shape(mpsz_pop2020) + tm_polygons ("AGED", style = "quantile", palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

## Selection Criteria
Instead of creating a facet map, the selection function can be used to map spatial objects meeting the selection criterion.
```{r}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ]) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          legend.hist = TRUE,
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.outside.size = 0.2,
            legend.outside.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```