---
title: "In-class Exercise 2"
author: "Fan Syong Yue"
date: 08/26/24
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
---

# Getting Started

We want to use `tidyverse` and `sf` packages

```{r}
pacman::p_load(tidyverse, sf)
```

# Working with 2014 Master Plan Planning Sub-zone Data

## Importing the data

We want to first import the `shp` shapefile

```{r}
mpsz14_shp <- st_read(dsn = "data/",
                        layer = "MP14_SUBZONE_WEB_PL")
```

Since the original downladed KML file errors out, let's create one using the SHP file

```{r}
#| output: false
st_write(mpsz14_shp,
         "data/MP14_SUBZONE_WEB_PL.kml",
         delete_dsn = TRUE)
```

We now want to import the `kml` file

```{r}
mpsz14_kml <- st_read("data/MP14_SUBZONE_WEB_PL.kml")
```

# Working with 2019 Master Plan Subzone Boundary (No Sea)

## Importing the data

We want to first import the `shp` shapefile

```{r}
mpsz19_shp <- st_read(dsn = "data/",
        layer = "MPSZ-2019")
```

We now want to import the `kml` file

```{r}
mpsz19_kml <- st_read("data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
```

# Handling Coordinate Systems
We want to transform our `mpsz19_shp` variable to use the EGS 3414 coordinate system instead of its original WGS 84
```{r}
mpsz19_shp_3414 <- st_transform(mpsz19_shp, crs = 3414)
st_crs(mpsz19_shp_3414)
```

We now want to import and transform our Preschool data into the same EGS 3414 coordinate system
```{r}
preschool_kml_3414 <- st_read("data/PreSchoolsLocation.kml") %>%
  st_transform(crs = 3414)
st_crs(preschool_kml_3414)
```
# Working with Population Data
We first need to import the population data into a data frame
```{r}
popdata <- read_csv("data/respopagesextod2023.csv")
```

We now want to break the data down into their planning area, subzone and age group
```{r}
popdata2023 <- popdata %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP`=sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from = AG, values_from = POP)

colnames(popdata2023)
```
## Data processing
We now want to process and split the data
```{r}
popdata2023 <- popdata2023 %>%
  mutate(`YOUNG` = rowSums(.[3:6]) + rowSums(.[14])) %>%
  mutate(`ECONOMY ACTIVE` = rowSums(.[7:13])+rowSums(.[15])) %>%
  mutate(`AGED` = rowSums(.[16:21])) %>%
  mutate(`TOTAL` = rowSums(.[3:21])) %>%
  mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, `ECONOMY ACTIVE`, `AGED`, `TOTAL`, `DEPENDENCY`)
```
Now we can join `popdata2023` and `mpsz19_shp_3414`
```{r}
popdata2023 <- popdata2023 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = list(toupper))
```

```{r}
mpsz_pop2023 <- left_join(mpsz19_shp_3414, popdata2023,
                          by = c("SUBZONE_N" = "SZ"))
glimpse(mpsz_pop2023)
```
```{r}
pop2023_mpsz <- left_join(popdata2023, mpsz19_shp_3414,
                          by = c("SZ" = "SUBZONE_N"))
glimpse(pop2023_mpsz)
```