---
title: "In-class Exercise 5"
author: "Fan Syong Yue"
date: 17/09/24
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
---

# Data

We will be using 2 data sets for this exercise

-   Hunan county boundary layer. This is a geospatial data set in ESRI shapefile format.
-   Hunan_2012.csv: This csv file contains selected Hunan’s local development indicators in 2012.

# Packages

We will be using the following packages:

-   sf
-   spdep
-   tmap
-   tidyverse
-   knitr
-   GWmodel

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, knitr, GWmodel)
```

# Data


```{r}
#| eval: false
hunan  <- st_read(dsn = "data/geospatial", layer = "Hunan")
```

```{r}
#| eval: false
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

## Joining the Data

We want to combine the data to update the `hunan` SpatialPolygonsDataFrame with the attribute fields of `hunan2012`

```{r}
#| eval: false
hunan_sf <- left_join(hunan,hunan2012)%>%
  select(1:3, 7, 15, 16, 31, 32)
```

We want to save the data that we have selected
```{r}
#| eval: false
write_rds(hunan_sf, "data/rds/hunan_sf.rds")
```

We can now check and reimport the data into `hunan_sf`

```{r}
hunan_sf <- read_rds("data/rds/hunan_sf.rds")
```

## Converting to SpatialPolygonDataFrame
```{r}
hunan_sp <- as_Spatial(hunan_sf)
```

# Geographically Weighted Summary Statistics
## Adaptive bandwith

### Cross Validation
```{r}
bw_CV <- bw.gwr(GDPPC ~ 1,
                 data = hunan_sp,
                 adaptive = TRUE,
                 kernel = "bisquare",
                 longlat = T)
```
### AIC
```{r}
bw_AIC <- bw.gwr(GDPPC ~ 1,
                 data = hunan_sp,
                 approach = "AIC",
                 adaptive = TRUE,
                 kernel = "bisquare",
                 longlat = T)
```

### Computing
```{r}
gwstat <- gwss(data = hunan_sp,
               vars = "GDPPC",
               bw = bw_AIC,
               adaptive = TRUE,
               kernel = "bisquare",
               longlat = T)
```

### Preparing output data
We want to extract the SDF data table from the gwss object and convert it into a data.frame by using `as.data.frame()`
```{r}
gwstat_df <- as.data.frame(gwstat$SDF)
```

We now want to append the derived data onto the `hunan_sf` data.frame. We will use `cbind()` to achieve this
```{r}
hunan_gstat <- cbind(hunan_sf, gwstat_df)
```
### Visualisation
We want to display the geographically weighted mean
```{r}
tm_shape(hunan_gstat) +
  tm_fill("GDPPC_LM",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Distribution of geograhpically weighted mean",
            main.title.position = "center",
            main.title.size = 2.0,
            legend.text.size = 1.2,
            legend.height = 1.50,
            legend.width = 1.50,
            frame = TRUE)
```



## Fixed Bandwith

### Cross Validation

```{r}
bw_fixed_CV <- bw.gwr(GDPPC ~ 1,
                 data = hunan_sp,
                 adaptive = FALSE,
                 kernel = "bisquare",
                 longlat = T)
```

### AIC
```{r}
bw_fixed_AIC <- bw.gwr(GDPPC ~ 1,
                 data = hunan_sp,
                 approach = "AIC",
                 adaptive = FALSE,
                 kernel = "bisquare",
                 longlat = T)
```

